A Dearth of Young Stars and Planets in the Kepler Field

### 2024-05-28 paper/ms.tex
update ms

The point of Figure~\ref{fig:agescalecompone} is hopefully clear:
until this study, the ages of the youngest Kepler stars and planets were not 
well-measured.
While previous studies reported accurate ages for
$\gtrsim$1\,Gyr stars, ages of sub-gigayear stars have historically been
overestimated by 0.3--2\,dex, with severely understated uncertainties.  
For isochrone and kinematic ages, this is because these
methods rely on parameters that do not appreciably change at
$t$$\lesssim$1\,Gyr.  For the rotation-based ages in
\citet{2023ApJ...952..131M} and \citet{2024AJ....167..159L}, the
discrepancy is
caused by details of the calibration methodologies.  The
\texttt{kiauhoku} spin-down model for instance has known
``stretches'' and ``compressions'' in its age scale relative to observed open clusters
\citep[see][Sec. 7.3]{2023ApJ...952..131M}.
The \citet{Mamajek_2008}
calibration used by \citet{Reinhold_2015} appears more accurate than the former
two models, because it 
was directly fit to reproduce the cluster sequences known at the time.  However the
uncertainties from all of these methods appear to be underestimated.  This
is probably because
they do not marginalize over
the range of rotation periods accessible to stars of a fixed
mass and age.

@@

The stellar ages 
are given in Table~\ref{tab:planets} for the known planet-hosts,
and in Table~\ref{tab:stars} for all Kepler stars.  Figure~\ref{fig:hist_tgyro} shows the rotation-based
age distributions of all the stars (left column) and the known
planet-hosts (middle column).  These ``histograms'' are sampled from
the posterior probability distributions by drawing ten random samples
from each posterior, and then computing the re-weighted
histogram of the resulting samples.  The uncertainties assume Poisson
statistics.  We truncated the plots near $4$\,Gyr since this is the
cluster age out to which our interpolation-based method is calibrated.
The overall slope yields a dearth of young stars relative to the expectation
of a uniform age distribution.
%In an alternative coinage, this tendency for fewer young
%stars to be present could also be dubbed a ``demographic cliff'',
%in reference to the phenomenon from human demography.

The age distributions of Kepler stars and KOI hosts in Figure~\ref{fig:hist_tgyro} appear
similar.  One diagnostic for whether the two
distributions are drawn from the same underlying distribution is the
Kolmogorov-Smirnov test;  we calculated this statistic in a manner
that accounts for the Poisson uncertainties by performing 1000 random
draws of 70\% of the stars from each sample.  The resulting 
range of $\log_{\rm 10}p$ (2.5$^{\rm th}$ to 97.5$^{\rm th}$ percentile)
spanned -3.9 to -1.2 for the K dwarfs,
and -5.2 to -2.2 for all Kepler stars.  This degree of similarity suggests 
that although weak differences may be present,
the ``common wisdom'' that young transiting planets are hard to detect
due to the variability of their host stars requires some
nuance.  If true, this statement only holds for the very youngest
stars, where we see a marginal deficit in the KOI host age
distribution relative to that of the overall stellar sample.
However at the age resolution of this study, 
this selection bias against sub-gigayear planets seems small.

An important separate selection effect concerns Kepler's ability to
detect the rotation signals.  \citet{2022ApJ...937...94M} studied this
issue, and found for Sun-like stars that the fraction of rotation
signals detectable by Kepler is near unity up to $\approx$3\,Gyr, and
that it drops to almost zero by $\approx$5\,Gyr .  This trend is
opposite in functional form to our derived age distributions.

>>>>

Figure~\ref{fig:agescalecompone} show that although
previous studies reported ages that agree with the cluster scale for
$\gtrsim$1\,Gyr stars, sub-gigayear stars have historically had their ages 
overestimated by 0.3--2\,dex, with severely understated uncertainties.  
For isochrone and kinematic ages, this is because these
methods rely on parameters that do not appreciably change at
$t$$\lesssim$1\,Gyr.  For the rotation-based ages in
\citet{2023ApJ...952..131M} and \citet{2024AJ....167..159L}, the
discrepancy is
caused by details of the calibration methodologies.  The
\texttt{kiauhoku} spin-down model for instance has known
``stretches'' and ``compressions'' in its age scale relative to observed open clusters
\citep[see][Sec. 7.3]{2023ApJ...952..131M}.
The \citet{Mamajek_2008}
calibration used by \citet{Reinhold_2015} appears more accurate than the former
two models, because it 
was directly fit to reproduce the cluster sequences known at the time.  However the
uncertainties from all of these methods appear to be underestimated.  This
is probably because
they do not marginalize over
the range of rotation periods accessible to stars of a fixed
mass and age.

@@

The stellar ages 
are given in Table~\ref{tab:planets} for the known planet hosts,
and in Table~\ref{tab:stars} for all Kepler stars.  Figure~\ref{fig:hist_tgyro} shows the rotation-based
age distributions of all the stars (left column) and the
KOI hosts (middle column).  These ``histograms'' are sampled from
the posterior probability distributions by drawing ten random samples
from each posterior, and then computing the re-weighted
histogram of the resulting samples.  The uncertainties assume Poisson
statistics.  We truncated the plots near $4$\,Gyr since this is the
cluster age out to which our interpolation-based method is calibrated.
The overall slope yields a dearth of young stars relative to the expectation
of a uniform age distribution.
%In an alternative coinage, this tendency for fewer young
%stars to be present could also be dubbed a ``demographic cliff'',
%in reference to the phenomenon from human demography.

The age distributions of Kepler stars and KOI hosts in Figure~\ref{fig:hist_tgyro} appear
similar.  One diagnostic for whether the two
distributions are drawn from the same underlying distribution is the
Kolmogorov-Smirnov test;  we calculated this statistic in a manner
that accounts for the Poisson uncertainties by performing 1000 random
draws of 70\% of the stars from each sample.  The resulting 
$\log_{\rm 10}p$ values spanned (2.5$^{\rm th}$ to 97.5$^{\rm th}$ percentiles)
-3.9 to -1.2 for the K dwarfs,
and -5.2 to -2.2 for all Kepler stars.  This suggests 
that while small differences may be significant, they are not drastic.
This similarity
contradicts
the common-sense argument that young transiting planets are hard to detect
due to the photometric variability of their host stars.  If true, this statement only holds for the very youngest
($<$0.4\,Gyr)
stars, where there is a marginal deficit in the KOI host age
distribution relative to that of the overall stellar sample.

An important separate selection effect concerns Kepler's ability to
detect the rotation signals.  \citet{2022ApJ...937...94M} studied this
issue, and found for Sun-like stars that the fraction of rotation
signals detectable by Kepler is near unity up to $\approx$3\,Gyr, and
that it drops to almost zero by $\approx$5\,Gyr.  This trend is
opposite in functional form to our derived age distributions.
